---
title: "introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide'}
library(tidyverse)
library(lubridate)
library(treeImpulses)
```

# Introduction to the treeImpulses package

There are two primary purposes of the treeImpulses package, 1) estimate tree-based impulse responses, and 2) plot tree-based impulse responses. Accordingly, this introduction will demonstrate 1) how to estimate tree-based impulse responses, and 2) how to plot tree-based impulse responses. 

## Standard impulse responses
### Prepare data 

First, prepare the data necessary to run the experiment. Here we will replicate a standard monetary VAR and will pull the required variables from FRED via the quantmod package.

```{r, results='hide'}

# real output gap and inflation
quantmod::getSymbols.FRED(c('GDPC1','GDPPOT','GDPDEF'), env = globalenv())
Y = 
  purrr::reduce(
    list(
      data.frame(GDPC1, date = zoo::index(GDPC1)),
      data.frame(GDPPOT, date = zoo::index(GDPPOT)),
      data.frame(GDPDEF, date = zoo::index(GDPDEF))
    ),
    dplyr::inner_join, by = 'date'
  ) %>% 
  dplyr::mutate(y = 100*((GDPC1-GDPPOT)/GDPPOT),
         p = 400*(GDPDEF - lag(GDPDEF))/lag(GDPDEF)) %>%
  dplyr::select(date, y, p)

# federal funds rate
quantmod::getSymbols.FRED(c('FEDFUNDS'), env = globalenv())
I = data.frame(FEDFUNDS, date = zoo::index(FEDFUNDS)) %>%
  dplyr::group_by(y = lubridate::year(date), q = lubridate::quarter(date)) %>%
  dplyr::summarize(date = min(date),
            ff = mean(FEDFUNDS, na.rm = T)) %>%
  dplyr::ungroup() %>%
  dplyr::select(date, ff) %>%
  data.frame()
  
# merge data
Data = 
  dplyr::inner_join(Y, I, by = 'date') %>%
  dplyr::filter(date < as.Date('2020-01-01')) %>%
  dplyr::arrange(date) %>%
  dplyr::select(-date) %>%
  na.omit()

# demean the data
Data = Data %>% 
  dplyr::mutate_all(function(X){return(X - mean(X, na.rm = TRUE))})

```

### Estimate tree-based impulse responses

Next, armed with a measure of inflation, the output gap, and federal funds rate, one may estimate impulse responses.

```{r, results='hide'}

# estimate impulse response  (using Trees)
impulseResponses = 
  localProjectionIRF_VAR(
    data = Data,                 
    shocks = c('y','p','ff'),                
    targets = c('y','p','ff'),
    horizons = 12,             # horizons to forecast out to
    lags = 4,                  # lags to include in regressions 
    lags_max = 12,             # maximum number of lags
    standardize = FALSE,       # standardize inputs (mean = 0 variance = 1)
    confidence = 0.1,          # alpha used for two-sided confidence intervals 
    states = NULL,             # column name used to define states
    NW = FALSE,                # Newey-West correction on variance-covariance matrix
    NW_lags = NULL,            # number of lags to use in Newey-West correction
    NW_prewhite = NULL,        # prewhite option for Newey-West correction (see sandwich::NeweyWest function)
    engine = 'AW',             # string declaring how to estimate responses 
    treeBag = 5,               # number of tree-based models to bag over
    cores = -1,                # number of cores available to use, set -1 to indicate use all available resources
    seed = 1116,               # random seed to be used (must be integer)
    IV = TRUE,                 # perform a two-stage instrumental variable regression (currently only supported for trees)
    Cholesky = FALSE           # append the initial structural shock from a cholesky decomposition [primarily for replication purposes]
  )

```

### Plot impulse responses

Lastly, one may plot the resulting impulse responses. 

```{r, fig.dim = c(6,6), fig.align='center'}
plot(impulseResponses)
```


## State-dependent impulse responses

Alternatively, one may want to impose explicit state-dependence in their impulse responses. To do this, include a factor or integer column in the design matrix, and declare this column as the states argument. 

```{r, results='hide'}

# create state variable
Data = Data %>% 
  dplyr::mutate(expansion = as.factor(dplyr::if_else(y >= 0, 'exp', 'rec')))

# estimate impulse response  
impulseResponses = 
  localProjectionIRF_VAR(
    data = Data,                 
    shocks = c('y','p','ff'),                
    targets = c('y','p','ff'),
    horizons = 12,             # horizons to forecast out to
    lags = 4,                  # lags to include in regressions 
    lags_max = 12,             # maximum number of lags
    standardize = FALSE,       # standardize inputs (mean = 0 variance = 1)
    confidence = 0.1,          # alpha used for two-sided confidence intervals 
    states = 'expansion',             # column name used to define states
    NW = FALSE,                # Newey-West correction on variance-covariance matrix
    NW_lags = NULL,            # number of lags to use in Newey-West correction
    NW_prewhite = NULL,        # prewhite option for Newey-West correction (see sandwich::NeweyWest function)
    engine = 'AW',             # string declaring how to estimate responses 
    treeBag = 5,               # number of tree-based models to bag over
    cores = -1,                # number of cores available to use, set -1 to indicate use all available resources
    seed = 1116,               # random seed to be used (must be integer)
    IV = TRUE,                 # perform a two-stage instrumental variable regression (currently only supported for trees)
    Cholesky = FALSE           # append the initial structural shock from a cholesky decomposition [primarily for replication purposes]
  )

```

And plot impulse responses by state.

```{r, fig.dim = c(6,6), fig.align='center'}
plot(impulseResponses, state = 'exp')
```


```{r, fig.dim = c(6,6), fig.align='center'}
plot(impulseResponses, state = 'rec')
```


### Conclusion

While the treeImpulses package focuses solely on two tasks, there is a great amount of flexibility in which one may accomplish that task, estimating tree-based impulse responses - the user may choose what type of random forest to use, choose how to identify the optimal number of lags, and generate instrumental variable based shocks. Please see function documentation for further details. 
